---
title: "Stat 341 -- Project Milestone 2"
author: "Gloria Grace & Sherise Immanuela"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    code_download: true
---

```{r, setup, include = FALSE, message=FALSE}
# load packages that are going to be used
library(fastR2)      # some data sets
library(pander)      # nicely formatted tables with pander()
library(knitr)       # so you can use kable() for tables
library(patchwork)   # for combining plots

# data wrangling
library(tidyverse)

# several packages for bayesian stuff -- more may come later
library(tidybayes)    
library(bayesplot)
library(dagitty)
library(CalvinBayes)

# Some customization. You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw())     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  tidy = FALSE,     # display code as typed (rather than reformatted)
  fig.width = 4,    # adjust this to make figures wider or narrower
  fig.height = 2.5, # adjust this to make figures taller or shorrter
  size = "small")   # slightly smaller font for code
```

```{r}
mlvar_data = read.csv("SA_data_for_mlvar_ana.csv")
#make subjects and dyads factors
mlvar_data$subject <- as.factor(mlvar_data$subject)
mlvar_data$dyad <- as.factor(mlvar_data$dyad)
#compute synchrony, log transform to correct for heteroscedasticity
mlvar_data$Synchrony <- -log(mlvar_data$dtw)
colnames(mlvar_data) <- c('X','dyad','subject','dtw','EyeContact','PupilSize','Synchrony')
```

```{r}
glimpse(mlvar_data)
```

The data is from a research conducted by Sophie Wohltjen and Thalia Wheatley. Their research paper can be found here: <https://www.pnas.org/doi/10.1073/pnas.2106645118>

This research paper examines the role of eye contact in conversation. The authors found that when two people are talking, eye contact is a key factor in determining when shared attention between them is rising or falling. It was found that when two people are engaging in conversation, eye contact increases as the conversation progresses, indicating a rise in shared attention.

For this project, the question we are interested in answering is: How does eye contact affect shared attention in conversations?

Our predictor variables are eye contact and pupil size the response variable is shared attention in conversation (synchrony).
Our predictor variables are eye contact and pupil size.
Our response variable is shared attention in conversation (synchrony).

```{r}
precision_ex <- dagitty("dag{
PupilSize -> Synchrony
Synchrony -> PupilSize
PupilSize -> EyeContact
EyeContact -> Synchrony
}")
gg_dag(precision_ex,
       size = 30,
       highlight = 'A')
```

$$\text{synchrony}_i \sim \text{Normal}(\mu_i, \sigma)$$

$$\mu_i = \beta_0 + \beta_1 * \text{EyeContact}_i + \beta_2 * \text{PupilSize}_i$$

$$\beta_0 \sim \text{Normal}(30,20)$$
$$\beta_1 \sim \text{Normal}(0.1, 0.05)$$

$$\beta_1 \sim \text{Normal}(0.1, 0.05)$$

$$\sigma \sim \text{Normal}(\text{mean}_2 = 10, \text{sd}_2 = 20)$$



